---
  output: github_document
---
  
  <!-- README.md is generated from README.Rmd. Please edit that file -->
  
  ```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# spatialrisk

spatialrisk is an R-package for spatial risk calculations. In particular, it can be used to determine concentration risk in the context of Solvency II. 

The package offers an effective approach to calculate the *standard formula* under Solvency II. The *standard formula* under Solvency II asks companies to report their largest fire concentration in respect of the fire peril within a radius of 200m. This is the maximum gross sum insured of the set of buildings fully or partly located within this radius. 



## Installation

You can install spatialrisk from github with:
  
  ```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("MHaringa/spatialrisk")
```

## Example 1
Find all observations in dataframe `Groningen` within radius of 50m from center point *(lon,lat) = (6.561561,53.21326)*:
  
  ```{r example}
library(spatialrisk)
points_in_circle(Groningen, lon_center = 6.571561, lat_center = 53.21326, radius = 50)
```

## Example 2
Find for each row in dataframe `df` the observations in dataframe `Groningen` within radius of 100m from lon/lat pair. Then the function takes for the obtained observations the sum of the column `amount`: 

```{r example2}
df <- data.frame(location = c("p1", "p2"), 
                 lon = c(6.561561, 6.561398), 
                 lat = c(53.21369, 53.21326))

concentration(df, Groningen, value = amount, radius = 100)
```





